Balance responses in MS

### 2020-01-30 R/msBalance.Rmd
first implementation of plotting timeseries data in R

# identify enrollment date for each participant and calculate amount of observation time
participants = participants %>% left_join(participants %>% group_by(Participant) %>% summarize(`Enrollment Date` = min(`Visit Date`, na.rm = TRUE))) %>% 
  mutate(`Time Since Enrollment (Days)` = ifelse(Group == "MS", `Visit Date` - `Enrollment Date`, 0))

@@

```{r}
# if trial-level data files do not exist, run matlab batch steps to populate them
if (!file.exists("../tmp/Participant.csv")|
    !file.exists("../tmp/Date.csv")|
    !file.exists("../tmp/Trial.csv")|
    !file.exists("../tmp/Time.csv")|
    !file.exists("../tmp/Treadmill.csv")|
    !file.exists("../tmp/Emg1.csv")|
    !file.exists("../tmp/Emg2.csv")|
    !file.exists("../tmp/Emg3.csv")|
    !file.exists("../tmp/Emg4.csv")){
  system(paste(matlabCommand, "'cd ../m; preprocessData; exit'"))
}
trialData = cbind(
vroom("../tmp/Participant.csv", col_names = FALSE) %>% select(Participant = X1),
vroom("../tmp/Date.csv", col_names = FALSE) %>% select(Date = X1),
vroom("../tmp/Trial.csv", col_names = FALSE) %>% select(Trial = X1)
) %>% as_tibble() %>% mutate(Date = ymd(Date))

time = vroom("../tmp/Time.csv", col_names = FALSE)
treadmill = vroom("../tmp/Treadmill.csv", col_names = FALSE)
emg1 = vroom("../tmp/Emg1.csv", col_names = FALSE)
emg2 = vroom("../tmp/Emg2.csv", col_names = FALSE)
emg3 = vroom("../tmp/Emg3.csv", col_names = FALSE)
emg4 = vroom("../tmp/Emg4.csv", col_names = FALSE)

@@

# the dataset emg1 corresponds to right TA. reconfigure it to enable tidy plotting.

@@

>>>>

# identify enrollment date for each participant and calculate amount of time
participants = participants %>% left_join(participants %>% group_by(Participant) %>% summarize(`Enrollment Date` = min(`Visit Date`, na.rm = TRUE))) %>% 
  mutate(`Time Since Enrollment (Days)` = ifelse(Group == "MS", `Visit Date` - `Enrollment Date`, 0))

@@

```{r}
# if trial-level data files do not exist, run matlab batch steps to populate them
if (!file.exists("../tmp/Participant.csv")|
    !file.exists("../tmp/Date.csv")|
    !file.exists("../tmp/Trial.csv")|
    !file.exists("../tmp/Time.csv")|
    !file.exists("../tmp/Treadmill.csv")|
    !file.exists("../tmp/Emg1.csv")|
    !file.exists("../tmp/Emg2.csv")|
    !file.exists("../tmp/Emg3.csv")|
    !file.exists("../tmp/Emg4.csv")){
  system(paste(matlabCommand, "'cd ../m; preprocessData; exit'"))
}

# code the trial-level information so that each is coded.
# this call also codes the groups and isolates first visits based on the subject code.
# finally, this code isolates the trial number and direction from the Trial code, _t1-_t6.
trialData = cbind(
  vroom("../tmp/Participant.csv", col_names = FALSE) %>% select(Participant = X1),
  vroom("../tmp/Date.csv", col_names = FALSE) %>% select(Date = X1),
  vroom("../tmp/Trial.csv", col_names = FALSE) %>% select(Trial = X1)
) %>% as_tibble() %>%
  mutate(Date = ymd(Date)) %>%
  mutate(`Code` = paste0("X", row_number())) %>% 
  mutate(`Study Group` = if_else(Participant %contains% "MSC", "Control", "MS")) %>% 
  mutate(Visit = if_else(Participant %contains% "F", "Follow-up", "Initial")) %>% 
  mutate(`Trial Number` = Trial %>% str_extract("_t\\d") %>% str_extract("\\d") %>% as.numeric()) %>% 
  mutate(`Perturbation Direction` = if_else(`Trial Number` <= 3, "Forward", "Backward"))

time = vroom("../tmp/Time.csv", col_names = FALSE) %>% select(Time = X1)
treadmill = vroom("../tmp/Treadmill.csv", col_names = FALSE)
emg1 = vroom("../tmp/Emg1.csv", col_names = FALSE)
emg2 = vroom("../tmp/Emg2.csv", col_names = FALSE)
emg3 = vroom("../tmp/Emg3.csv", col_names = FALSE)
emg4 = vroom("../tmp/Emg4.csv", col_names = FALSE)

@@

# the dataset emg1 corresponds to right TA. reconfigure it to enable tidy plotting.
taR = time %>%  cbind(emg1) %>% as_tibble() %>% tidyr::pivot_longer(-Time, names_to = "Code", values_to = "V") %>%
  mutate(Muscle = "Tibialis Anterior") %>% mutate(Side = "Right") %>% select(`Code`, Muscle, Side, everything())

# I cannot figure out how to get dplyr::left_join to join by a variable with backticks, so rename temporarily
taR = left_join(taR, trialData, by = "Code") %>% 
  select(`Study Group`,Participant, Date, Trial, `Code`, `Trial Number`, `Perturbation Direction`, Muscle, Side, everything()) %>% 
  arrange(`Study Group`, Participant, `Trial Number`, Time)

# note that a large amount of the data (40%) is follow-up, with different visit amounts. for now, let's just consider the initial visit.
taR = taR %>% filter(Visit == "Initial")

# consider only those participants for whom both some EMG and minimal demographic data are available.
includedParticipants = intersect(taR$Participant, participants$Participant)
taR = taR %>% filter(Participant %in% includedParticipants)
participants = participants %>% filter(Participant %in% includedParticipants)

@@

```{r rawEmg}
library(ggplot2)
p = ggplot(taR, aes(x = Time, y = V, group = Code)) + geom_line() + facet_grid(`Study Group` ~ `Perturbation Direction`)
p
```
